---
title: 'Plotting Part A: Temperatures'
author: "Lilian Marie Boll"
date: "08/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
```

## Plot Secundary Structure over 6 different Temperatures

This script is part of the project for the subject MSI of the Master Bioinformatics for Health Science at the UPF, 3rd trimester 2021.
The following script generates a dataframe for each temperature that must be provided as a txt file. The output is a plot containing the percentage of alpha helix over time.

The generated plot shows the percentage of alpha-helices over time for each temperature:

```{r}
temp <- 5
c_5 <- read.table('pcts_278K.txt') %>%
  data.table::transpose() %>% # Have a column per timepoint
  select(V1:V201) %>%
  `colnames<-`(0:200) %>% # Add time to column names
  pivot_longer(cols=colnames(.), names_to="time", values_to="ss") %>% # Change the table layout
  group_by(time) %>% # Perform the counting grouping by the different times
  count(ss) %>% # Count the different types of SS elements for each time group
  filter(grepl("H", ss)) %>% # Filter only the rows with the helices counts
  mutate(pct = n/10, # Compute the helix percentage (nº of residues with helix over the total 10)
         time =  as.numeric(time), # Make the time column numeric
         temp =  temp) %>% # Add the temperature for plotting
  select(time, pct, temp)

temp <- 25
c_25 <- read.table('pcts_298K.txt') %>%
  data.table::transpose() %>% # Have a column per timepoint
  select(V1:V201) %>%
  `colnames<-`(0:200) %>% # Add time to column names
  pivot_longer(cols=colnames(.), names_to="time", values_to="ss") %>% # Change the table layout
  group_by(time) %>% # Perform the counting grouping by the different times
  count(ss) %>% # Count the different types of SS elements for each time group
  filter(grepl("H", ss)) %>% # Filter only the rows with the helices counts
  mutate(pct = n/10, # Compute the helix percentage (nº of residues with helix over the total 10)
         time =  as.numeric(time), # Make the time column numeric
         temp =  temp) %>% # Add the temperature for plotting
  select(time, pct, temp)

temp <- 50
c_50 <- read.table('pcts_323K.txt') %>%
  data.table::transpose() %>% # Have a column per timepoint
  select(V1:V201) %>%
  `colnames<-`(0:200) %>% # Add time to column names
  pivot_longer(cols=colnames(.), names_to="time", values_to="ss") %>% # Change the table layout
  group_by(time) %>% # Perform the counting grouping by the different times
  count(ss) %>% # Count the different types of SS elements for each time group
  filter(grepl("H", ss)) %>% # Filter only the rows with the helices counts
  mutate(pct = n/10, # Compute the helix percentage (nº of residues with helix over the total 10)
         time =  as.numeric(time), # Make the time column numeric
         temp =  temp) %>% # Add the temperature for plotting
  select(time, pct, temp)


temp <- 100
c_100 <- read.table('namd_100/mailinglist.txt') %>%
  data.table::transpose() %>% # Have a column per timepoint
  select(V1:V201) %>%
  `colnames<-`(0:200) %>% # Add time to column names
  pivot_longer(cols=colnames(.), names_to="time", values_to="ss") %>% # Change the table layout
  group_by(time) %>% # Perform the counting grouping by the different times
  count(ss) %>% # Count the different types of SS elements for each time group
  filter(grepl("H", ss)) %>% # Filter only the rows with the helices counts
  mutate(pct = n/10, # Compute the helix percentage (nº of residues with helix over the total 10)
         time =  as.numeric(time), # Make the time column numeric
         temp =  temp) %>% # Add the temperature for plotting
  select(time, pct, temp)


temp <- 150
c_150 <- read.table('namd_150/mailinglist.txt') %>%
  data.table::transpose() %>% # Have a column per timepoint
  select(V1:V201) %>%
  `colnames<-`(0:200) %>% # Add time to column names
  pivot_longer(cols=colnames(.), names_to="time", values_to="ss") %>% # Change the table layout
  group_by(time) %>% # Perform the counting grouping by the different times
  count(ss) %>% # Count the different types of SS elements for each time group
  filter(grepl("H", ss)) %>% # Filter only the rows with the helices counts
  mutate(pct = n/10, # Compute the helix percentage (nº of residues with helix over the total 10)
         time =  as.numeric(time), # Make the time column numeric
         temp =  temp) %>% # Add the temperature for plotting
  select(time, pct, temp)


temp <- 200
c_200 <- read.table('namd_200/mailinglist.txt') %>%
  data.table::transpose() %>% # Have a column per timepoint
  select(V1:V201) %>%
  `colnames<-`(0:200) %>% # Add time to column names
  pivot_longer(cols=colnames(.), names_to="time", values_to="ss") %>% # Change the table layout
  group_by(time) %>% # Perform the counting grouping by the different times
  count(ss) %>% # Count the different types of SS elements for each time group
  filter(grepl("H", ss)) %>% # Filter only the rows with the helices counts
  mutate(pct = n/10, # Compute the helix percentage (nº of residues with helix over the total 10)
         time =  as.numeric(time), # Make the time column numeric
         temp =  temp) %>% # Add the temperature for plotting
  select(time, pct, temp)

c_5$temp <-  "5ºC"
c_25$temp <-  "25ºC"
c_50$temp <-  "50ºC"
c_100$temp <-  "100ºC"
c_150$temp <-  "150ºC"
c_200$temp <-  "200ºC"
data_all <- rbind(c_5, c_25, c_50, c_100, c_150, c_200)
head(data_all)

```

```{r}
ggplot(transform(data_all,
                 temperature=factor(temp,levels=c("5ºC", "25ºC", "50ºC", "100ºC", "150ºC", "200ºC")))) + 
  geom_line(aes(time,pct))+ facet_grid(temperature~.) +
  ggtitle("Effect of temperature on % of alpha-helix, simulated in vacuum") +
  xlab("Time (ps)") + ylab("% of alpha-helix") +
  coord_cartesian(ylim = c(0.4, 0.9)) +
  theme(axis.title=element_text(size=14,face="bold")) +
  theme_bw() +
  ggsave(filename = "temps_vacuum.png", width = 6.5, height = 5.5, dpi=2000)
```


# Plot % at timeframe 150

```{r plot, echo=FALSE}
#Water dataframe
percentages<- c(0.8,0.8,0.8,0.7,0,0)
temperatures<- c(5, 25, 50, 100, 150, 200)
data_150<- as.data.frame(cbind(percentages,temperatures))
data_150$environment<-"Water"

## Vacuum dataframe
percentages <- c(0.9,0.9,0.9,0.8,0.6,0.7)
temperatures<- c(5, 25, 50, 100, 150, 200)
data_francho<- as.data.frame(cbind(percentages,temperatures))
data_francho$environment <- "Vacuum"

## Join dataframes
picoseconds_all <- rbind(data_150, data_francho)       

## Plot
ggplot(transform(picoseconds_all,
                 environment=factor(environment,levels=c("Water","Vacuum")))) + 
  geom_line(aes(temperatures,percentages, group=1))+ facet_grid(environment~.) +
  theme_bw()+
  labs(title = "Effect of temperature on % of alpha-helix in different environments", 
       subtitle = "At 150 ps") +
  theme(title = element_text(face = "bold"), plot.subtitle = element_text(face = "italic"))+
  xlab("Temperature (ºC)") + ylab("% of alpha-helix") +
  theme(axis.title=element_text(size=14,face="bold"))
```